noam_beta_diversity

### 2024-08-26 noam_beta_diversity.Rmd
Ran metaMDS on bray curtis matrices at habitat level and on full data. Created dissimilarity over time (interval) model and graph.

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "taxa",
  values_to = "percent_cover")

 transect_level <- long_data %>%
   dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
   dplyr::summarise(mean_cover = mean(percent_cover), 
             SE = sd(percent_cover)/sqrt(n()))

site_level <- long_data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
  dplyr::summarise(mean_site_cover = mean(percent_cover), 
            site_se = sd(percent_cover)/sqrt(n()))

# subset based on habitat
fringing <- subset(data, habitat == "fringing")
backreef <- subset(data, habitat == "backreef")
outer10 <- subset(data, habitat == "outer_10")
outer17 <- subset(data, habitat == "outer_17")

@@

# bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
# summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
# 
# bray_mod_reduced <- glmmTMB(dissimilarity_trans ~ time*habitat + time*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series)

@@

```{r dissimilarity over time plot, fig.height=15, fig.width=15}
 ggplot(bray_time_series,
        aes(x = time, y = dissimilarity, colour = island_side) )+
   facet_wrap(~habitat) + 
   geom_point() +
   scale_colour_manual(values = island_side_colours) +
   stat_smooth(se = T, linewidth = 2) +
   labs(y = "Dissimilarity\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

```{r fringing nmds set up}
# identifying species w/ > 10% cover for plotting later

#### FRINGING
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
# fringing_columns_to_keep <- apply(fringing_species_matrix, 2, function(x) any(x >= 0.1))
# fringing_abundant_species_matrix <- as.data.frame(fringing_species_matrix)[ , fringing_columns_to_keep]
# fringing_abundant_species_names <- colnames(fringing_abundant_species_matrix)
# 
# fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
# 
# # for mean centroid:
# fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
# 
# # takes several hours to run - do so at your own peril
# set.seed(1032) 
# # ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
# fringing_ord3 # stress = 0.018
# 
# 
# sppscores(fringing_ord3) <- fringing_species_matrix # add back in species scores
# fringing.species.scores <- as.data.frame(scores(fringing_ord3, "species")) # some NaN for absent species
# fringing.species.scores <- na.omit(fringing.species.scores)
# fringing.species.scores$species <- rownames(fringing.species.scores) 
# 
# 
# fringing.site.scores <- as.data.frame(vegan::scores(fringing_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# fringing.site.scores$stratum <- rownames(fringing.site.scores)
# fringing.site.scores <- merge(fringing_subset_meta, fringing.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(site.scores, here("data", "fringing_site_scores.csv"))
# write_csv(species.scores, here("data", "fringing_species_scores.csv"))

@@

fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
# fringing.cent.meta <- fringing_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# fringing.cent <- merge(fringing.cent, fringing_subset_meta[,-which(names(fringing_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]

#############
# BACKREEF #
############
backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)

backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]

# for mean centroid:
backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
backreef_ord3 # stress = 0.018

sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
backreef.species.scores <- na.omit(backreef.species.scores)
backreef.species.scores$species <- rownames(backreef.species.scores)

backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
backreef.site.scores$stratum <- rownames(backreef.site.scores)
backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")

write_csv(site.scores, here("data", "backreef_site_scores.csv"))
write_csv(species.scores, here("data", "backreef_species_scores.csv"))

@@

outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)

outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]

# for mean centroid:
outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
outer10_ord3 # stress = 0.018

sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
outer10.species.scores <- na.omit(outer10.species.scores)
outer10.species.scores$species <- rownames(outer10.species.scores)

outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer10.site.scores$stratum <- rownames(outer10.site.scores)
outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer10_site_scores.csv"))
write_csv(species.scores, here("data", "outer10_species_scores.csv"))

@@

outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)

outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]

# for mean centroid:
outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
outer17_ord3 # stress = 0.018

sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
outer17.species.scores <- na.omit(outer17.species.scores)
outer17.species.scores$species <- rownames(outer17.species.scores)

outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer17.site.scores$stratum <- rownames(outer17.site.scores)
outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer17_site_scores.csv"))
write_csv(species.scores, here("data", "outer17_species_scores.csv"))

@@

```{r full bray curtis set up, eval = FALSE}
### I DON'T THINK ANY COMPUTER CAN RUN THIS
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year,"-", data$location, sep = "")
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year,"-", data_subset$location, sep = "") # for merging with bray matrix later
brayfull <- vegdist(species_matrix, "bray") # create bray dissimilarity matrix
# View(brayfull)
braymat <- as.matrix(brayfull) # conver to matrix object
# convert to pairwise comparison
comp <- braymat[lower.tri(braymat)] # extract lower triangle
row_indices <- row(braymat)[lower.tri(braymat)] # extract row indices for first quad in comparison
col_indices <- col(braymat)[lower.tri(braymat)] # 
loc1 <- rownames(braymat)[row_indices]
loc2 <- colnames(braymat)[col_indices]
pairwise_df <- data.frame(loc1 = loc1, loc2 = loc2, dissimilarity = comp)
# pairwise_df$year1 <- sub("(.*)-.*", "\\1", pairwise_df$loc1)
year1 <- strsplit(pairwise_df$loc1, "-")[[1]]

>>>>

#pivot long 
# long_data <- pivot_longer(
#   data = data,
#   cols = 8:77,
#   names_to = "taxa",
#   values_to = "percent_cover") 
# 
# 
#  transect_level <- long_data %>%
#    dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
#    dplyr::summarise(mean_cover = mean(percent_cover), 
#              SE = sd(percent_cover)/sqrt(n()))
#  
# site_level <- long_data %>%
#   dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
#   dplyr::summarise(mean_site_cover = mean(percent_cover), 
#             site_se = sd(percent_cover)/sqrt(n()))
# 
# subset based on habitat
fringing <- subset(data, habitat == "fringing")
backreef <- subset(data, habitat == "backreef")
outer10 <- subset(data, habitat == "outer_10")
outer17 <- subset(data, habitat == "outer_17")

@@

bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
# model.sel(dredge(bray_mod_universal)) # top model as specified
summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
hist(residuals(bray_mod_universal))
plot(predict(bray_mod_universal) ~ residuals(bray_mod_universal)) # looks like reflection of logit curve

emtrends(bray_mod_universal, pairwise ~ habitat |island_side, "time")
bray_trends_emm <- emmip(bray_mod_universal, habitat ~ time | island_side, at = list(time = seq(1,16,1)), plotit = F, CIs = T)
inv_logit <- function(x) {
  exp(x)/(1+exp(x))
}
```

```{r dissimilarity over time plot, fig.height=15, fig.width=15}
 ggplot()+
   geom_point(data = bray_time_series, aes(x = time, y = dissimilarity, colour = island_side), alpha = 0.01 ) +
   geom_line(data = bray_trends_emm, aes(x = time, y = inv_logit(yvar), colour = island_side), linewidth = 2) +
   geom_ribbon(data = bray_trends_emm, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = island_side), alpha = 0.3) +
   facet_wrap(~habitat) +

 #  scale_colour_manual(values = island_side_colours) +
   stat_smooth(se = T, linewidth = 2) +
   labs(y = "Dissimilarity\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

@@

```{r hab level nmds set up}

#### WARNING: THIS CHUNK TAKES ~2 DAYS TO RUN! SUGGESTEDF TO USE PROVIDED CSV FILES WITH ORDINATION DATA ######

#### FRINGING

# fringing_columns_to_keep <- apply(fringing_species_matrix, 2, function(x) any(x >= 0.1))
# fringing_abundant_species_matrix <- as.data.frame(fringing_species_matrix)[ , fringing_columns_to_keep]
# fringing_abundant_species_names <- colnames(fringing_abundant_species_matrix)
# 
# fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
# 
# # for mean centroid:
# fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
#

# set.seed(1032) 
# # fringing_ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
# fringing_ord3 # stress = 0.018
# 
# 
# sppscores(fringing_ord3) <- fringing_species_matrix # add back in species scores
# fringing.species.scores <- as.data.frame(scores(fringing_ord3, "species")) # some NaN for absent species
# fringing.species.scores <- na.omit(fringing.species.scores)
# fringing.species.scores$species <- rownames(fringing.species.scores) 
# 
# 
# fringing.site.scores <- as.data.frame(vegan::scores(fringing_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# fringing.site.scores$stratum <- rownames(fringing.site.scores)
# fringing.site.scores <- merge(fringing_subset_meta, fringing.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(site.scores, here("data", "fringing_site_scores.csv"))
# write_csv(species.scores, here("data", "fringing_species_scores.csv"))

@@

# fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
# fringing.cent.meta <- fringing_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# fringing.cent <- merge(fringing.cent, fringing_subset_meta[,-which(names(fringing_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
# fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]

#############
# BACKREEF #
############
# backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
# backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
# backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)
# 
# backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]
# 
# # for mean centroid:
# backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")
# 
# # takes several hours to run - do so at your own peril
# set.seed(1032) 
# backreef_ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
# backreef_ord3 # 
# 
# 
# sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
# backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
# backreef.species.scores <- na.omit(backreef.species.scores)
# backreef.species.scores$species <- rownames(backreef.species.scores) 
# 
# 
# backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# backreef.site.scores$stratum <- rownames(backreef.site.scores)
# backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(backreef.site.scores, here("data", "backreef_site_scores.csv"))
# write_csv(backreef.species.scores, here("data", "backreef_species_scores.csv"))

@@

# outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
# outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
# outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)
# 
# outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]
# 
# # for mean centroid:
# outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")
#

# set.seed(1032) 
# outer10_ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
# outer10_ord3 # 
# 
# 
# sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
# outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
# outer10.species.scores <- na.omit(outer10.species.scores)
# outer10.species.scores$species <- rownames(outer10.species.scores) 
# 
# 
# outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# outer10.site.scores$stratum <- rownames(outer10.site.scores)
# outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(outer10.site.scores, here("data", "outer10_site_scores.csv"))
# write_csv(outer10.species.scores, here("data", "outer10_species_scores.csv"))

@@

## outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
## outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
## outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)
## 
## outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]
## 
## # for mean centroid:
## outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")
##

## set.seed(1032) 
## outer17_ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
## outer17_ord3 # 
## 
## 
## sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
## outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
## outer17.species.scores <- na.omit(outer17.species.scores)
## outer17.species.scores$species <- rownames(outer17.species.scores) 
## 
## 
## outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
## # add back in metadata
## outer17.site.scores$stratum <- rownames(outer17.site.scores)
## outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")
## 
## 
## 
## 
## write_csv(outer17.site.scores, here("data", "outer17_site_scores.csv"))
## write_csv(outer17.species.scores, here("data", "outer17_species_scores.csv"))

@@

```{r full ordination}
### TAKES ~1 wk to run 
# full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
# full_species_matrix <- full_species_matrix/100 # convert from percent to proportion
# rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
# full_rows_to_remove <- rowSums(full_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
# full_species_matrix <- full_species_matrix[!full_rows_to_remove,] 
# full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
# full_subset <- data[!full_rows_to_remove,] # make sure rows are also removed from dataframe
# full_subset$stratum <- paste(full_subset$year,"-", full_subset$location, sep = "") # for merging with bray matrix later
# brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
# # View(brayfull)
# full_braymat <- as.matrix(brayfull) # conver to matrix object
# 
# full_columns_to_keep <- apply(full_species_matrix, 2, function(x) any(x >= 0.1))
# full_abundant_species_matrix <- as.data.frame(full_species_matrix)[ , full_columns_to_keep]
# full_abundant_species_names <- colnames(full_abundant_species_matrix)
# 
# full_subset_meta <- full_subset[ , which(names(full_subset) %in% meta_cols)]
# 
# # for mean centroid:
# full_subset_meta$island_side_year <- paste(full_subset_meta$year, "-", full_subset_meta$island_side_habitat, sep = "")
# 
# # takes several hours to run - do so at your own peril
# set.seed(1032) 
# full_ord3 <- metaMDS(full_braymat, dist	= "bray", k = 3, trymax = 100) 
# full_ord3 # 
#

# sppscores(full_ord3) <- full_species_matrix # add back in species scores
# full.species.scores <- as.data.frame(scores(full_ord3, "species")) # some NaN for absent species
# full.species.scores <- na.omit(full.species.scores)
# full.species.scores$species <- rownames(full.species.scores) 
# 
# 
# full.site.scores <- as.data.frame(vegan::scores(full_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# full.site.scores$stratum <- rownames(full.site.scores)
# full.site.scores <- merge(full_subset_meta, full.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(full.site.scores, here("data", "full_site_scores.csv"))
# write_csv(full.species.scores, here("data", "full_species_scores.csv"))
